---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1 (Chapter 3, #3, 6 marks)
Given:
Predictors: $X_1$=GPA,  $X_2$=IQ,  $X_3$=Gender(1 for Female, 0 for Male),  $X_4$=Interaction between GPA and IQ,  $X_5$=Interaction between GPA and Gender <br/>
Response: Starting Salary after graduation(in thousands of dollars) <br/>
Coefficient Estimates: $\hat{\beta_0}$=50, $\hat{\beta_1}$=20, $\hat{\beta_2}$=0.07, $\hat{\beta_3}$=35, $\hat{\beta_4}$=0.01, $\hat{\beta_5}$=-10 <br/>

(a) Based on information given in question, using least squares to fit the model, the relationship between response and predictors can be respresented by the following equation: <br/>
Let response be represented by Y. Therefore, prediction for starting salary after graduation will be $\hat{Y}$ <br/>

$\hat{Y}$ = $\hat{\beta_0}$ + $\hat{\beta_1}$$X_1$ + $\hat{\beta_2}$$X_2$ + $\hat{\beta_3}$$X_3$ + $\hat{\beta_4}$$X_4$ + $\hat{\beta_5}$$X_5$ <br/>

Substituting values for predictors and coefficient estimates as given in question, we get: <br/>
$\hat{Y}$ = 50 + (20 x GPA) + (0.07 x IQ) + (35 x Gender) + (0.01 x Interaction between GPA and IQ) + 
((-10) x Interaction between GPA and Gender) <br/>

Interaction between a and b can be represented by their product. Thus, we get <br/>
$\hat{Y}$ = 50 + (20 x GPA) + (0.07 x IQ) + (35 x Gender) + (0.01 x GPA x IQ) + 
((-10) x GPA x Gender) <br/>

Substituting Gender=1, we get prediction for starting salary for Females: <br/>
$\hat{Y}_f$ = 50 + (20 x GPA) + (0.07 x IQ) + 35 + (0.01 x GPA x IQ) + (-10 x GPA) <br/>
$\hat{Y}_f$ = 85 + (10 x GPA) + (0.07 x IQ) + (0.01 x GPA x IQ) <br/>

Similarly, prediction for starting salary for males(Gender=0) is represented by: <br/>
$\hat{Y}_m$ = 50 + (20 x GPA) + (0.07 x IQ) + (0.01 x GPA x IQ) <br/>

Since the value of IQ and GPA is fixed(IQ x GPA is fixed), so prediction for starting salary for both females and males depends on GPA. <br/>

85 + (10 x GPA) $\leq$ 50 + (20 x GPA), if GPA is high enough <br/>

Thus, For a fixed value of IQ and GPA, males earn more on average than females provided that the GPA is high enough(Option 3)
<br/>

(b) As seen in part(a), prediction for starting salary for females(Gender=1) is represented by: <br/>
$\hat{Y}_f$ = 85 + (10 x GPA) + (0.07 x IQ) + (0.01 x GPA x IQ) <br/>
Given: IQ=110 and GPA=4, we get <br/>
$\hat{Y}_f$ = 85 + (10 x 4) + (0.07 x 110) + (0.01 x 4 x 110) <br/>
$\hat{Y}_f$ = 85 + 40 + 7.7 + 4.4 <br/>
$\hat{Y}_f$ = 137.1 <br/>

Thus, for given values of IQ and GPA, the predicted starting salary for females after graduation is 137.1(in thousand dollars) <br/>

(c) False (WE NEED TO FIND P_VALUES I GUESS)

## Question 2 (Chapter 3, #9, 10 marks)

```{r}
library(ISLR) 
data(Auto)
library(dplyr)
Auto <- 
  Auto %>% select(-name) %>% mutate(origin = factor(origin))
head(Auto)
attach(Auto)
```

(a) Scatterplot matrix which includes all of the variables in the data set is given by:
```{r}
pairs(Auto)
```

(b) Matrix of correlations between the variables(except name variable, which is qualitative) using the function cor():
```{r}
cor(Auto[1:7])
```
WHAT ABOUT ORIGIN IN THIS CASE, IT'S ALSO QUALITATIVE

(c) Multiple linear regression with mpg as the response and all other variables except name as the predictors: 
```{r}
lm.fit = lm(mpg~., data=Auto)
summary(lm.fit)
```
i) <br/>

ii) Predictors: Displacement, weight, year and origin appear to have a statistically significant relationship to the response. <br/>

iii) The year variable suggests that there is an increase of --- mpg per year. <br/>

(d)
```{r}
par(mfrow = c(2, 2))
plot(lm.fit)
```

(e)

(f) To keep the investigation of transformations
manageable, try transformations of the `weight` variable
only.

```{r}
par(mfrow = c(2, 2))
plot(log(weight), mpg)
plot(sqrt(weight), mpg)
plot((weight)^2, mpg)
```

## Question 3 (Chapter 4, #4, 7 marks)

(a)

(b)

(c)

(d)

(e)


## Question 4 (Chapter 4, #10 parts (a)-(h), 9 marks)

(a)

(b)

(c)

(d)

(e)

(f)

(g)

(h)

(i) DO NOT HAND IN THIS PART (though you are, of course,
free to do it on your own).