---
  title: "Homework 1"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 (Chapter 2, #1, 8 marks)

(a) A flexible model is better. Since we have large sample size and small number of predictors, we can fit the model closer to data and have better performance without overfitting.

(b) A flexible model is worse. As the number of observations is very small and number of predictors is extremely large, a flexible model would have problems of high variance due to over-fitting and fitting outliers.

(c) A flexible model would fit better. If the relationship between the predictors and response is highly non-linear, a flexible model will be less restrictive on the shape of fit. With more degrees of freedom, a flexible method would fit better.

(d) A flexible model is worse. A flexible model will further increase the variance by fitting to noise terms. 


## Question 2 (Chapter 2, #2, 6 marks)

(a)
* Regression  # predicting a continuous quantity output (CEO salary)  
* Inference   # determining relationship between predictors and output(CEO salary)
* n = 500 observations #collected data belongs to 500 firms in US
* p = 3 variables
* profit
* number of employees
* industry

(b) 
* Classification # predicting discrete class labels, success/failure
* Prediction  # interested only about outcomes(success or failure) and not the relationship
* n = 20 observations  #data for 20 similar products are collected
* p = 13 variables  
* price charged for products
* marketing budget
* competition price
* ten other variables


(c) 
* Regression  #the result (% change in the US Dollar) is quantitative
* Prediction  #interested only in % change of USD
* n = 52 weekly observations  #No of weeks in the year 2012
* p = 3 variables
* % change in the US market
* % change in the British market
* % change in the German market

## Question 3 (Chapter 2, #9, 8 marks)

```{r}
#install.packages("ISLR")
library(ISLR)
data(Auto)
head(Auto)
autodata <- na.omit(Auto) #used to get rid of the missing values
attach(autodata) #tells R to make the variables in this data frame available by name
```
<br/>
  
  (a) Quantitative Predictors: mpg, cylinders, displacement, horsepower, weight, acceleration, year
Qualitative Predictors: Origin(needs to be factored as qualitative), name

* Cylinders can also be treated as qualitative predictor.
<br/>
  
  (b) The range of each quantitative predictor can be given by: 
  ```{r}
sapply(autodata[,1:7],range) #it applies range function over all seven quantitative predictors where first row signifies minimum values of each predictor and the second row shows the maximum values of predictors 
```
<br/>
  
  (c) The mean and standard deviation of each quantitative predictor is given by:
  ```{r}
meansd <- function(x) {
  out<-c(mean(x),sd(x))
  names(out) <- c("mean","SD")
  out
}

apply(autodata[,1:7],MARGIN=2,FUN=meansd)
```
<br/>
  
  (d) 10th through 85th observations can be removed using:
  ```{r}
AutoSubset <- autodata[-c(10:85),c(1:7)]
```

The range, mean, and standard deviation of each predictor in the subset of the data that remains:
  ```{r}
sapply(AutoSubset[,1:7],range) #first row in the output signifies minimum values of each predictor and the second row shows the maximum values of predictors 
sapply(AutoSubset[,1:7],mean)
sapply(AutoSubset[,1:7],sd)
```

<br/>
  
  (e) Relationship among predictors: 
  
  Some plots depicting relationship between predictors are shown below:
  ```{r}
pairs(autodata) #creates a scatterplot for every pair of predictors in Auto data set
plot(cylinders,mpg)
plot(mpg,displacement)
plot(mpg,horsepower)
plot(mpg,weight)
plot(year,mpg)

plot(cylinders,displacement)
plot(cylinders,acceleration)

as.factor(origin)
boxplot(origin,mpg)
```

* mpg is inversely related to displacement, horsepower and weight, however, it increases as year increases.
* mpg is maximum for 4 cylinders as compared to others. mpg is inversely related to cylinders.
* cylinders is positively related to displacement, horsepower and weight but negatively related to acceleration.

<br/>
  
  (f) Correlation Matrix representing correlation between predictors:
  ```{r}
#install.packages("GGally")
library("ggplot2")
library("GGally") #using GGally to plot correlation matrix
ggcorr(autodata, label=TRUE) # command to plot correlation matrix
```

From the graphical plots of pairs(autodata) and correlation matrix we can say that: <br/>
  displacement and weight are highly correlated to horsepower.
<br/> cylinders, horsepower, year and origin can be used as strong predictors to predict gas mileage(mpg).
